Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-02 GECCO/volunteer-clouds-gecco16.tex
Minor corrections and comments for conclusions

In the second implementation, Web Workers
were used so that several clients per browser page could be executed and
run in the background; our intention was to increase the time spent by
the users and, in fact, the number of generations per
user increased to the point that we obtained a performance better than
baseline in most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment simultaneously, so that our objective was
achieved. In general, one of the first conclusions for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved using web workers.

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model and its parameters depends on the implementation, with contributions following
a GEV distribution with single-thread and a
Weibull distribution when web workers are used. These distributions are
close enough to each other and, in the second case, reflect the fact
that most time spent in the page is actually devoted to computing, as
opposed to presenting the chart or not doing anything,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse implication, from
games to volunteer computing,  might also be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they have participated
once, they would be encouraged to do for a longer time or to share
it. In the spirit of Open Science, this 
gamification might involve computing in real-time data such as the one
presented in this paper, while showing it in the same page.

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. % (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
% Moved to conclusions. 
There are several validity threats to the design of this study.
Participants of the experiments were mainly members of our 
own social networks, personal colleagues, friends 
and acquaintances, therefore to extend this work other options 
should be considered in order to include 
a wider audience with different motives for participation.
Another threat is that participants were anonymous so the amount of repeated 
participations and the effects this might have had in the outcome is not
known. Another issue is that the scalability of the system was not thoroughly tested
but for the amount of participants the basic configuration 
was more than adequate, then again more experiments are needed in order
to assess the capability both with more clients and other server
configurations. Although care was taken to include representative benchmarks
of real world problems, more experiments with a broader coverage 
should allow a better understanding of the practical limitations of
the architecture and the tools chosen.
%  Even if
% the three previous experiments were published in a similar way, one
% obtained up to 5 times more total cycles  than the one with the least
% number of cycles. 
It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation

@@

% Finally, the implementation needs some refinement in terms of
% programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
% to create a generator in which the user just has to create a fitness
% function, with the rest of the framework wrapped around
% automatically.
%  All this, as well as the data used for this paper and the paper
% itself, has been published with a free license in GitHub at
% \url{https://github.com/JJ/modeling-volunteer-computing}.

>>>>

In the second implementation, Web Workers
were used so that several clients per browser page could be executed and
run in the background; our intention was to increase the time spent by
the users and, in fact, the number of generations per
user increased to the point that we obtained a performance better than
baseline in most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment simultaneously, so that our objective was
achieved. In general, one of the first conclusions of this paper is
that the browser technologies should be used to its full extent, so
that the user time is leveraged to its full potential, and this was
achieved using web workers.

The second objective of this paper was to model the user behavior in a
first attempt to try and predict performance. As should be expected,
the model and its parameters depend on the implementation, with contributions following
a GEV distribution with single-thread and a
Weibull distribution when web workers are used. These distributions are
close enough to each other and, in the second case, reflect the fact
that most time spent in the page is actually devoted to computing, as
opposed to presenting the chart or not doing anything,
which is why the time spent (represented by the number of
contributions to the pool) follows a model quite similar to that found
for games or other online activities. The reverse implication, from
games to volunteer computing,  might also be true: if we
want to have returning users for the experiments, it is probable that
we should {\em gamify} the experience so that once they have participated
once, they would be encouraged to do for a longer time or to share
it. In the spirit of Open Science, this 
gamification might involve computing in real-time data such as the one
presented in this paper, while showing it in the same page.

In general, linking and finding correlations between user choices and
performance is an interesting avenue to explore in the future. % (Paloma) Mario, isn't this the future work? Should it be in conclusions? If not, just ignore this comment.
% Moved to conclusions. 
There are several validity threats to the design of this study.
Participants of the experiments were mainly members of our 
own social networks, personal colleagues, friends 
and acquaintances, therefore, to extend this work other options 
should be considered in order to include 
a wider audience with different motives for participation.
Another threat is that participants were anonymous, so the amount of repeated 
participations and the effects this might have had in the outcome is not
known. Another issue is that the scalability of the system was not thoroughly tested
but for the amount of participants the basic configuration 
was more than adequate, then again more experiments are needed in order
to assess the capability both with more clients and other server
configurations. Although care was taken to include representative benchmarks
of real world problems, more experiments with a broader coverage 
should allow a better understanding of the practical limitations of
the architecture and the tools chosen.
%  Even if
% the three previous experiments were published in a similar way, one
% obtained up to 5 times more total cycles  than the one with the least
% number of cycles. 
It is also essential to obtain volunteers as
simultaneously as possible, so it is possible that the features of the
social network in terms of real-time use will also play a big
role. Even as it is difficult to create controlled experiments in this
area, it is an interesting challenge to face in the future. % (Paloma) Wrong collocation

@@

% Finally, the implementation needs some refinement in terms of
% programming and also ease of use. Tools such as Yeoman might be used % ¿añadir una cita o la URL http://yeoman.io/ a pie de página ?
% to create a generator in which the user just has to create a fitness
% function, with the rest of the framework wrapped around
% automatically.
%  All this, as well as the data used for this paper and the paper
% itself, has been published with a free license in GitHub at
% \url{https://github.com/JJ/modeling-volunteer-computing}.  
% Pablo: I should uncomment this last paragraph, but using the url: https://github.com/ANONYMOUS
